---
title: "R Notebook"
output: html_notebook
---


# Disclaimer
- this code is forked from https://www.kaggle.com/pranav84/talkingdata-eda-to-model-evaluation-lb-0-9683/data

```{r libraries}
if (!require("pacman")) install.packages("pacman")

pacman::p_load(knitr, tidyverse, highcharter, data.table, lubridate, pROC, tictoc, DescTools, lightgbm, caret)
set.seed(84)               
options(scipen = 9999, warn = -1, digits= 5)

```

# Paramters
```{r parameters}
train_file = "../data/train.csv"
test_file= "../data/test.csv"
nrows_to_test = 100000
nrows_to_train = 10000000
output_file = "../submit/lm_R_6mil_v1.csv"
is_test = TRUE

```


```{r load data}
tic("Total time for processing")
if(is_test){
        training = fread(train_file, nrows = nrows_to_test,
                      col.names = c("ip", "app", "device", "os", "channel", "click_time",
                                    "attributed_time", "is_attributed"), ,
                      showProgress = FALSE) 
}else{
        training = fread(train_file, nrows = nrows_to_train,
                      col.names = c("ip", "app", "device", "os", "channel", "click_time",
                                    "attributed_time", "is_attributed"), ,
                      showProgress = FALSE) 
        
}

training = training %>%
        select(-attributed_time)%>%
        mutate(wday = Weekday(click_time), hour = hour(click_time)) %>%
        mutate_all(as.factor) %>%
        select(-click_time) %>%
        add_count(ip, wday, hour) %>% rename("nip_day_h" = n) %>%
        add_count(ip, hour, channel) %>% rename("nip_h_chan" = n) %>%
        add_count(ip, hour, os) %>% rename("nip_h_osr" = n) %>%
        add_count(ip, hour, app) %>% rename("nip_h_app" = n) %>%
        add_count(ip, hour, device) %>% rename("nip_h_dev" = n) %>%
        select(-ip)

toc()
head(training)
```



```{r load test}
tic("Total time for processing")
if(is_test){
        testing = fread(test_file, showProgress = FALSE, nrows = nrows_to_test) 
}else{
        testing = fread(test_file, showProgress = FALSE)
}

# extract click_id for submission file
sub = data.table(click_id = test$click_id, is_attributed = NA)

testing = testing %>%
        mutate(wday = Weekday(click_time), hour = hour(click_time)) %>%
        mutate_all(as.factor) %>%
        select(-click_time) %>%
        add_count(ip, wday, hour) %>% rename("nip_day_h" = n) %>%
        add_count(ip, hour, channel) %>% rename("nip_h_chan" = n) %>%
        add_count(ip, hour, os) %>% rename("nip_h_osr" = n) %>%
        add_count(ip, hour, app) %>% rename("nip_h_app" = n) %>%
        add_count(ip, hour, device) %>% rename("nip_h_dev" = n) %>%
        select(-c(ip))

toc()

```

# Data Analysis
```{r eda}
str(training)

```

```{r kable}
?kable
kable(as.data.frame(lapply(training, function(x) length(unique(x)))))

kable(as.data.frame(lapply(testing, function(x) length(unique(x)))))
```

## Categorical features
h1 = train %>% group_by(app) %>% summarise(count = n()) %>%
        arrange(desc(count)) %>% head(15) %>% mutate(app = as.character(app)) %>%
        hchart("bar", hcaes(x=app, y = count, color = -count)) %>%
        hc_add_theme(hc_theme_ffx()) %>% hc_title(text = "Top Apps")

h2 <- train %>% group_by(os) %>% summarise(count = n()) %>% 
  arrange(desc(count)) %>% head(15) %>% mutate(os = as.character(os)) %>%
  hchart("bar", hcaes(x = os, y = count, color =-count)) %>%
  hc_add_theme(hc_theme_ffx()) %>% hc_title(text = "Top OS")

h3 <- train %>% group_by(channel) %>% summarise(count = n()) %>% 
  arrange(desc(count)) %>% head(15) %>% mutate(channel = as.character(channel)) %>%
  hchart("bar", hcaes(x = channel, y = count, color =-count)) %>%
  hc_add_theme(hc_theme_ffx()) %>% hc_title(text = "Top Channels")

h4 <- train %>% group_by(device) %>% summarise(count = n()) %>% 
  arrange(desc(count)) %>% head(15) %>% mutate(device = as.character(device)) %>%
  hchart("bar", hcaes(x = device, y = count, color =-count)) %>%
  hc_add_theme(hc_theme_ffx()) %>% hc_title(text = "Top Devices")

h5 <- train %>% group_by(hour) %>% summarise(count = n()) %>%
  arrange(desc(count)) %>% head(15) %>% mutate(hour = as.character(hour)) %>%
  hchart("bar", hcaes(x = hour, y = count, color =-count)) %>%
  hc_add_theme(hc_theme_ffx()) %>% hc_title(text = "Top Hours")

hw_grid(h1, h2, h3, h4, h5, ncol = 5, rowheight = 600)


```{r}
?hchart

```


# Test Data
h1 <- test %>% group_by(app) %>% summarise(count = n()) %>% 
  arrange(desc(count)) %>% head(15) %>% mutate(app = as.character(app)) %>%
  hchart("bar", hcaes(x = app, y = count, color =-count)) %>%
  hc_add_theme(hc_theme_ffx()) %>% hc_title(text = "Top Apps")

h2 <- test %>% group_by(os) %>% summarise(count = n()) %>% 
  arrange(desc(count)) %>% head(15) %>% mutate(os = as.character(os)) %>%
  hchart("bar", hcaes(x = os, y = count, color =-count)) %>%
  hc_add_theme(hc_theme_ffx()) %>% hc_title(text = "Top OS")

h3 <- test %>% group_by(channel) %>% summarise(count = n()) %>% 
  arrange(desc(count)) %>% head(15) %>% mutate(channel = as.character(channel)) %>%
  hchart("bar", hcaes(x = channel, y = count, color =-count)) %>%
  hc_add_theme(hc_theme_ffx()) %>% hc_title(text = "Top Channels")

h4 <- test %>% group_by(device) %>% summarise(count = n()) %>% 
  arrange(desc(count)) %>% head(15) %>% mutate(device = as.character(device)) %>%
  hchart("bar", hcaes(x = device, y = count, color =-count)) %>%
  hc_add_theme(hc_theme_ffx()) %>% hc_title(text = "Top Devices")

h5 <- test %>% group_by(hour) %>% summarise(count = n()) %>%
  arrange(desc(count)) %>% head(15) %>% mutate(hour = as.character(hour)) %>%
  hchart("bar", hcaes(x = hour, y = count, color =-count)) %>%
  hc_add_theme(hc_theme_ffx()) %>% hc_title(text = "Top Hours")

h6 <- train %>% group_by(wday) %>% summarise(count = n()) %>%
  arrange(desc(count)) %>% mutate(wday = as.character(wday)) %>%
  hchart("bar", hcaes(x = wday, y = count, color =-count)) %>%
  hc_add_theme(hc_theme_ffx()) %>% hc_title(text = "Top Weekdays")

hw_grid(h1, h2, h3, h4, h5, ncol = 5, rowheight = 600)




# Modelling 
## test train split
```{r caret split}
set.seed(3456)
trainIndex = createDataPartition(training$is_attributed, p = .95, list = F)
dtrain = training[trainIndex, ]
valid = training[-trainIndex, ]

cat("train size : ", dim(dtrain), " | valid size : ", dim(valid), " | test  size : ", dim(testing))
```

# Preprocessing
```{r preprocess}


```


```{r table}
table(dtrain$is_attributed)
```

## Umbalancedness
```{r umbalance}
scale_pos_weight = 100 - ((220/100000)*100) # this will default in 50, as 50% of the data is positive
scale_pos_weight = 1/0.002

dtrain$weights = ifelse(dtrain$is_attributed == 0, 1, scale_pos_weight)

head(dtrain)
```


## Categorical data
- light gbm uses these without breaking into one-hot coding

categorical_features = c("app", "device", "os", "channel", "wday", "hour")

dtrain = lgb.Dataset(data = as.matrix(dtrain[, colnames(dtrain)!= "is_attributed"]),
                     label = dtrain$is_attributed, 
                     categorical_feature = categorical_features)

dvalid = lgb.Dataset(data = as.matrix(valid[, colnames(valid)!= "is_attributed"]),
                     label = valid$is_attributed, 
                     categorical_feature = categorical_features)

## trainControl
```{r trainControl}
fitControl = trainControl(##10-fold CV,
        method = "repeatedcv",
        number = 2,
        repeats = 2)

?gbm

?train
```



## GBM
```{r gbm}
tic("time for model tuning")
gbmFit1 = train(is_attributed ~ ., data = dtrain,
                method = "gbm",
                metric = "ROC",
                weights = weights,
                tuneLength = 30,
                trControl = fitControl,
                verbose = F)
toc()
```


# Model evaluation
## Feature importance
```{r feature importance}
fi = lgb.importance(model, percentage = TRUE)

highchart() %>%
        hc_title(text = "Feature importance by Gain") %>%
        hc_xAxis(categories = fi$Feature) %>%
        hc_add_series(name = "Gain", data = fi, type = "bar", hcaes(x = Feature, y = Gain)) %>%
        hc_add_theme(hc_theme_ffx())

```

## Validation
```{r ROC}
valid = head(valid, 50000)
val_preds = predict(model, data = as.matrix(valid[, colnames(valid)!="is_attributed"]), n = model$best_iter)


roc(valid$is_attributed, val_preds, plot = TRUE, col = "steelblue", lwd = 3,
    levels = base::levels(as.factor(valid$is_attributed)), grid = TRUE)

```


# Individual predictions

```{r individual predictions}
valid[1,]
```

```{r predicted value}
paste("predicted value", val_preds[[1]])

```

```{r reasoning}
tree_interpretation = lgb.interprete(model, data = as.matrix(valid[, colnames(valid)]), 1)
lgb.plot.interpretation(tree_interpretation[[1]])

```

# Submittion
```{r submit}
preds = predict(model, data = as.matrix(test[, colnames(test)], n = model$best_iter))
preds = as.data.frame(preds)
sub$is_attributed = preds
fwrite(sub, output_file)
```


